### 2021-05-12 manuscript.Rmd
Fix typos.

* __Background__: Multidrug-resistant organisms (MDROs) colonizing the healthcare environment have been shown to contribute to risk for healthcare-associated infections (HAIs), with adverse effects on patient morbidity and mortality. We sought to define <span style="background-color:lightcyan">how bacterial contamination and MDRO persistence in the healthcare environment are related to the position of patients and wastewater sites.</span>

@@

* __16S rRNA Gene Sequencing, <span style='background-color:lightcyan'>Quantitative PCR,</span> and Bioinformatics__: <span style='background-color:lightcyan'>A subset of samples were included for microbiome analysis, representing the earliest samples collected and processed prior to the COVID-19 pandemic.</span> Parallel surface samples were collected with flocked nylon swabs 16S rRNA gene sequencing and microbiome analysis. Swabs were frozen at -80C prior to DNA extraction and sequencing. DNA extraction was performed using MoBio PowerSoil DNA isolation kit. The V1–V2 hypervariable region of the 16S rRNA gene was amplified using barcoded primers 27F (5′AATGATACGGCGACCACCGAGATCTACACTATGGTAATTGTAGAGTTTGATCCTGGCTCAG-3′) and 338R (5′CAAGCAGAAGACGGCATACGAGATNNNNNNNNNNNNAGTCAGTCAGCCTGCTGCCTCCCGTAGGAGT-3′). Thermocycler conditions were as follows: 5 min at 95 °C, 30*(30 s at 95 °C, 30 s at 56 °C, 90 s at 72 °C), 8 min at 72 °C. 16S rRNA gene sequencing was performed with the Illumina MiSeq platform with 250-bp paired-end reads, permitting high-quality coverage of the 316-bp V1–V2 amplicon. Amplicon sequence variants (ASVs) were identified using DADA2 within the Quantitative Insights Into Microbial Ecology 2 (QIIME2) pipeline [@Callahan2019-ir; @Bolyen2019-xk]. Genus-level taxonomic assignment was performed by comparison to the Greengenes reference database [@McDonald2012-tf], using the naive Bayes classifier implemented in scikit-bio [@Bokulich2018-xf]. <span style='background-color:lightcyan'>To account for the compositional nature of microbiome measures, which may confound the analysis of relationships between bacterial community members,</span> the centered log-ratio transform was applied to genus-level abundances prior to modeling [@Gloor2017-kj]. <span style='background-color:lightcyan'>16S rRNA gene quantitative PCR (qPCR) also targeted the V1-V2 hypervariable region, with primers BSF8 (5’-AGAGTTTGATCCTGGCTCAG-3’) and BSR357 (5’-CTGCTGCCTYCCGTA-3’), and probe 5’-TAACAC ATGCAAGTCGA-3’. The standard curve was generated with a plasmid containing the 16S rRNA gene derived from _Streptococcus_.</span>

@@

* __Impact of Distance from Patient on Bacterial DNA Deposition__: We were able to complement comprehensive, culture-based surveillance with 16S rRNA gene sequencing to detect bacterial DNA deposition in the healthcare environment in a subset of 41 specimens from 16 of our cohort's subjects. In **Figure 2**, we compare the impact of distance from patient on the bacterial DNA reads assigned to the 11 genera from which MDROs were cultured. We found that _Enterobaceriaceae_, including _Citrobacter_, _Enterobacter_, _Klebsiella_, and _Raoultella_, exhibited a similar pattern as did bacterial cultures, increasing with distance from the patient and in proximity to the wastewater sites. In contrast, bacterial DNA assigned to _Acinetobacter_ and _Pseudomonas_ was distributed more evenly across all areas, though bacterial culture was more likely to detect these genera in the wastewater sites. Bacterial DNA reads assigned to _Staphylococcus_ were also distributed evenly across all areas, though bacterial culture was more likely to detect MRSA in close proximity to patients.

* __Relationship Between Bacterial DNA Deposition and MDRO Detection in the Healthcare Environment__: To evaluate the relationship between bacterial DNA contamination detected by untargeted 16S rRNA gene sequencing and environmental MDRO risk, we performed mixed-effects regression to determine the association between genus-level bacterial DNA abundance and the probability of MDRO detection across all observed MDRO genera and sampling sites. A positive association was observed across multiple genera and sites, but the only association that achieved high posterior certainty was _Pseudomonas_ in the far-from-patient site, <span style='background-color:lightcyan'>finding that as the probability of detecting _Pseudomonas_ DNA increased, so too did the probability of culturing CR _Pseudomonas_ at the site </span>(OR 1.47, 95%CI 1.09 - 1.99). <span style='background-color:lightcyan'>We further examined how bacterial community features, defined by 16S rRNA gene sequencing and qPCR, changed with distance from the patient and wastewater sites, as well as the association between bacterial community features and MDRO detection. We found that total bacterial abundance (16S rRNA gene copies per cm2 of surface area measured by qPCR) increased with proximity to wastewater sites, but the increase was small (1466 copies per cm2, 95%CI 670 – 2454 copies per cm2). Bacterial community Shannon diversity and richness (number of ASVs) did not change significantly with distance (**Figure 3**). The only association between a bacterial community feature (i.e., total bactertial abundance, diversity or richness) and probability of MDRO detection that achieved high posterior certainty was the association between total 16S rRNA gene copies and _Pseudomonas_ detection (OR 1.19, 95%CI 1.03 – 1.41). </span>

@@

<span style='background-color:lightcyan'>We compared culturable MDROS to bacterial DNA deposition in order to identify areas where bacteria are deposited but are not viable due to either environmental features (e.g., microenvironments that do not support bacterial growth or persistence) or cleaning practices. </span>In some cases, the pattern of culturable MDROs matched bacterial DNA deposition: _Enterobacterales_ detection increased with distance from the patient and in proximity to the wastewater sites, <span style='background-color:lightcyan'> mirroring the overall detection of 16S copies or bacterial abundance</span>. But in other cases, the discrepancy between bacterial DNA detection and culturable MDROs suggested environmental selection: though bacterial DNA assigned to _Pseudomonas_ and _Staphylococcus_ was distributed evenly across all areas, bacterial culture identified distinct hotspots for CR _Pseudomonas_ and MRSA, far from and near patients, respectively. Of note, both bacterial DNA assigned to _Enterococcus_ and culturable VRE were detected across distances, suggesting that bacterial deposition from stool is extensive. The decreasing probability of _C. difficile_ and decreasing _Clostridium_ DNA deposition as sampling became closer to the wastewater sites may be attributable to more effective cleaning practices specific to these pathogens within wastewater sites.

>>>>

* __Background__: Multidrug-resistant organisms (MDROs) colonizing the healthcare environment have been shown to contribute to risk for healthcare-associated infections (HAIs), with adverse effects on patient morbidity and mortality. We sought to define <span style="background-color:lightcyan">how bacterial contamination and persistent MDRO colonization in the healthcare environment are related to the position of patients and wastewater sites.</span>

@@

* __16S rRNA Gene Sequencing, <span style='background-color:lightcyan'>Quantitative PCR,</span> and Bioinformatics__: <span style='background-color:lightcyan'>A subset of samples (123 from 16 subjects) were included for microbiome analysis, representing the earliest samples collected and processed prior to the COVID-19 pandemic.</span> Parallel surface samples were collected with flocked nylon swabs 16S rRNA gene sequencing and microbiome analysis. Swabs were frozen at -80C prior to DNA extraction and sequencing. DNA extraction was performed using MoBio PowerSoil DNA isolation kit. The V1–V2 hypervariable region of the 16S rRNA gene was amplified using barcoded primers 27F (5′AATGATACGGCGACCACCGAGATCTACACTATGGTAATTGTAGAGTTTGATCCTGGCTCAG-3′) and 338R (5′CAAGCAGAAGACGGCATACGAGATNNNNNNNNNNNNAGTCAGTCAGCCTGCTGCCTCCCGTAGGAGT-3′). Thermocycler conditions were as follows: 5 min at 95 °C, 30*(30 s at 95 °C, 30 s at 56 °C, 90 s at 72 °C), 8 min at 72 °C. 16S rRNA gene sequencing was performed with the Illumina MiSeq platform with 250-bp paired-end reads, permitting high-quality coverage of the 316-bp V1–V2 amplicon. Amplicon sequence variants (ASVs) were identified using DADA2 within the Quantitative Insights Into Microbial Ecology 2 (QIIME2) pipeline [@Callahan2019-ir; @Bolyen2019-xk]. Genus-level taxonomic assignment was performed by comparison to the Greengenes reference database [@McDonald2012-tf], using the naive Bayes classifier implemented in scikit-bio [@Bokulich2018-xf]. <span style='background-color:lightcyan'>To account for the compositional nature of microbiome measures, which may confound the analysis of relationships between bacterial community members,</span> the centered log-ratio transform was applied to genus-level abundances prior to modeling [@Gloor2017-kj]. <span style='background-color:lightcyan'>16S rRNA gene quantitative PCR (qPCR) also targeted the V1-V2 hypervariable region, with primers BSF8 (5’-AGAGTTTGATCCTGGCTCAG-3’) and BSR357 (5’-CTGCTGCCTYCCGTA-3’), and probe 5’-TAACAC ATGCAAGTCGA-3’. The standard curve was generated with a plasmid containing the 16S rRNA gene derived from _Streptococcus_.</span>

@@

* __Impact of Distance from Patient on Bacterial DNA Deposition__: We were able to complement comprehensive, culture-based surveillance with 16S rRNA gene sequencing to detect bacterial DNA deposition in the healthcare environment in a subset of 123 specimens from 16 of our cohort's subjects. In **Figure 2**, we compare the impact of distance from patient on the bacterial DNA reads assigned to the 11 genera from which MDROs were cultured. We found that _Enterobaceriaceae_, including _Citrobacter_, _Enterobacter_, _Klebsiella_, and _Raoultella_, exhibited a similar pattern as did bacterial cultures, increasing with distance from the patient and in proximity to the wastewater sites. In contrast, bacterial DNA assigned to _Acinetobacter_ and _Pseudomonas_ was distributed more evenly across all areas, though bacterial culture was more likely to detect these genera in the wastewater sites. Bacterial DNA reads assigned to _Staphylococcus_ were also distributed evenly across all areas, though bacterial culture was more likely to detect MRSA in close proximity to patients.

* __Relationship Between Bacterial DNA Deposition and MDRO Detection in the Healthcare Environment__: To evaluate the relationship between bacterial DNA contamination detected by untargeted 16S rRNA gene sequencing and environmental MDRO risk, we performed mixed-effects regression to determine the association between genus-level bacterial DNA abundance and the probability of MDRO detection across all observed MDRO genera and sampling sites. A positive association was observed across multiple genera and sites, but the only association that achieved high posterior certainty was _Pseudomonas_ in the far-from-patient site, <span style='background-color:lightcyan'>finding that as the probability of detecting _Pseudomonas_ DNA increased, so too did the probability of culturing CR _Pseudomonas_ at the site </span>(OR 1.47, 95%CI 1.09 - 1.99). <span style='background-color:lightcyan'>We further examined how bacterial community features, defined by 16S rRNA gene sequencing and qPCR, changed with distance from the patient and wastewater sites, as well as the association between bacterial community features and MDRO detection. We found that total bacterial abundance (16S rRNA gene copies per cm2 of surface area measured by qPCR) increased with proximity to wastewater sites, but the increase was small (1466 copies per cm2, 95%CI 670 – 2454 copies per cm2). Bacterial community Shannon diversity and richness (number of ASVs) did not change significantly with distance (**Figure 3**). The only association between a bacterial community feature (i.e., total bacterial abundance, diversity or richness) and probability of MDRO detection that achieved high posterior certainty was the association between total 16S rRNA gene copies and _Pseudomonas_ detection (OR 1.19, 95%CI 1.03 – 1.41). </span>

@@

<span style='background-color:lightcyan'>We compared culturable MDROs to bacterial DNA deposition in order to identify areas where bacteria are deposited but are not viable due to either environmental features (e.g., microenvironments that do not support bacterial growth or persistence) or cleaning practices. </span>In some cases, the pattern of culturable MDROs matched bacterial DNA deposition: _Enterobacterales_ detection increased with distance from the patient and in proximity to the wastewater sites, <span style='background-color:lightcyan'> mirroring the overall detection of 16S copies or bacterial abundance</span>. But in other cases, the discrepancy between bacterial DNA detection and culturable MDROs suggested environmental selection: though bacterial DNA assigned to _Pseudomonas_ and _Staphylococcus_ was distributed evenly across all areas, bacterial culture identified distinct hotspots for CR _Pseudomonas_ and MRSA, far from and near patients, respectively. Of note, both bacterial DNA assigned to _Enterococcus_ and culturable VRE were detected across distances, suggesting that bacterial deposition from stool is extensive. The decreasing probability of _C. difficile_ and decreasing _Clostridium_ DNA deposition as sampling became closer to the wastewater sites may be attributable to more effective cleaning practices specific to these pathogens within wastewater sites.
